\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-13 papers/itsc_2012/final.tex
referencing figures and reducing wasted vspace

\usepackage[pdftex]{graphicx}
% TODO remember we need to meet color compliance:
% http://its.papercept.net/conferences/support/general.php
\newcommand{\pix}[3]{
  \begin{figure}[h]
    \centering \includegraphics[#3]{#1}
    \caption{#2}
  \end{figure}
}

@@

\begin{abstract} 
Autonomous vehicles have seen great advancements in recent years and
such vehicles are now ever closer to being commercially available.
The advent of driver-less cars provides opportunities for optimizing
traffic in ways not possible before. This paper introduces an open
source multi-agent traffic simulator called AORTA, which stands for
\textit{Approximately Orchestrated Road Traffic Automata}, and
demonstrates its usage for optimizing autonomous traffic at a
city-wide scale. AORTA creates scale simulations by generating maps
using publicly available road data from \textit{OpenStreetMap}
(OSM). Simulations of specific regions throughout the world can be set
up in the few minutes it takes to export data for the region of
interest from OSM. AORTA focuses on interactions between agents by
defining \textit{behaviors} for individual vehicles and
intersections. These behaviors allow agents to interact with each
other as well as with an external \textit{overseer} that is
responsible for optimizing traffic flow. Overseers can interact with
these agents to allow for dynamic route re-planning to reduce traffic
load, optimized intersection navigation, and improved car following.
\commentpk{rest devoted to application}
\end{abstract}

@@

\pix{downtown_atx.png}{A simulation of downtown Austin, Texas.}{scale=0.25}

@@

\commentd{consider re-grouping sections... maybe this is all 'methodology'?} \commentpk{I think it is fine as it is}

% Diagrma made with gliffy.com. architecture.gxml should enable editing.
\pix{architecture.png}
    {An illustration of AORTA's components.}
    {scale=0.18}

AORTA is divided into three modular components: the map model, micro-simulation
engine, and user interface (UI). The map model transforms OSM maps into AORTA graphs, then
answers pathfinding and geometry queries. The simulation engine adds a notion of
agents, vehicle dynamics, and collisions. Finally, the UI can
interactively render the map and agents. A headless mode also enables running
experiments without the overhead of visualization.

@@

\commentpk{need to reference the architecture figure here. the text in the
figure also needs to be bigger}

@@

Likewise, intersections check for collisions by examining the turn each agent is
performing and verifying that no other agent is simultaneously performing a
conflicting turn. AORTA has a low-granularity model of turn conflict. If two
vehicles at any point along two different turns could ever come into contact,
then the turns are always in conflict, no matter where the agents actually are
along that turn. The alternative of tiling intersections \cite{JAIR08-dresner}
would provide more granularity, but it adds computational complexity and would
be impractical for many of the complex intersection geometries inferred from
OSM.

@@

\pix{turn_conflicts.png}
    {An agent cannot turn left while agents on the perpendicular road cross}
    {scale=0.5}

@@

\pix{gridlock.png}{Each of the red agents wants to turn left, so none of them
can. The traffic jam will slowly build up around this intersection.}{scale=0.25}

% TODO in fact, this is caused by lack of lane changing!!!
% but that's scary... bad routes really do cause it.

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. AORTA can detect this by finding cycles in the
graph of the ``is following'' relation.

@@

Each policy has parameters, so the configuration used is relevant. Stop signs
accept agents FCFS, refusing agents until they have idled at $0$ speed for $1$
second. This delay approximates human reaction time. The traffic signal policy
requires timings and groups of simultaneous turns for each intersection. The
configuration tested here uses a breadth-first search from one major road's
lanes to try to synchronize green signals for the agents near that lane to not
stop for several intersections. This heuristic needs improvement in how it
handles the remaining groups of turns that are scheduled after the first
sequence. Finally, the autonomous reservation policy allows new agents to join
existing groups of agents if their turn is compatible with the others. To
prevent one heavy direction of traffic from hogging the intersection
indefinitely, a timeout preempts and cycles through reservations. Ideally this
should be based on the relative numbers of agents depending on their
reservation.

@@

\pix{avg_lag_agent_atx.png}
    {The distribution of average delay reveals that traffic signals can cause the most lag.}
    {scale=0.35}

If an intersection immediately permits an agent to proceed and grants other
agents in the same lane similar access (because this new agent must still follow
these others), then the new agent may simply proceed at full speed limit without
pausing. This gives the least possible delay spent along some lane, so any extra
time is defined as lag and is used to measure intersection performance. On
average, reservation policies cause less than $1$ second of delay at each
intersection, stop signs cause about $7$ seconds lag, and traffic signals cause
$16$ seconds.

\pix{agent_cnt_atx.png}
    {Autonomous reservations outperform stop signs, and traffic signal gridlocked.}
    {scale=0.35}

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. As one
measure of intersection policy efficiency, it reveals that autonomous
reservations are twice as efficient as the competition, with a steady-state of
about $400$ agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.

>>>>

\usepackage[pdftex]{graphicx}
% TODO remember we need to meet color compliance:
% http://its.papercept.net/conferences/support/general.php

@@

\begin{abstract} 
Autonomous vehicles have seen great advancements in recent years, and
such vehicles are now ever closer to being commercially available.
The advent of driver-less cars provides opportunities for optimizing
traffic in ways not possible before. This paper introduces an open
source multi-agent traffic simulator called AORTA, which stands for
\textit{Approximately Orchestrated Road Traffic Automata}, and
demonstrates its usage for optimizing autonomous traffic at a
city-wide scale. AORTA creates scale simulations by generating maps
using publicly available road data from \textit{OpenStreetMap}
(OSM). Simulations of specific regions throughout the world can be set
up in the few minutes it takes to export data for the region of
interest from OSM. AORTA focuses on interactions between agents by
defining \textit{behaviors} for individual vehicles and
intersections. These behaviors allow agents to interact with each
other as well as with an external \textit{overseer} that is
responsible for optimizing traffic flow. Overseers can interact with
these agents to allow for dynamic route re-planning to reduce traffic
load, optimized intersection navigation, and improved car following.
\commentpk{rest devoted to application}
\end{abstract}

@@

\begin{figure}[h]
  \centering \includegraphics[width=\linewidth]{downtown_atx.png}
  \caption{A simulation of downtown Austin, Texas.}
  \label{fig:ui_screenshot}
  \vspace{-10pt}
\end{figure}

% TODO reference the screenshot

@@

% Diagram made with gliffy.com. architecture.gxml should enable editing.
\begin{figure*}
  \centering \includegraphics[scale=0.3]{architecture.png}
  \caption{A summary of AORTA's components.}
  \label{fig:arch}
  \vspace{-20pt}
\end{figure*}

AORTA is divided into three modular components: the map model, micro-simulation
engine, and user interface (UI). The map model transforms OSM maps into AORTA graphs, then
answers pathfinding and geometry queries. The simulation engine adds a notion of
agents, vehicle dynamics, and collisions. Finally, the UI can
interactively render the map and agents. A headless mode also enables running
experiments without the overhead of visualization. The interaction between
modules is visualized in figure \ref{fig:arch}.

@@

@@

Likewise, intersections check for collisions by examining the turn each agent is
performing and verifying that no other agent is simultaneously performing a
conflicting turn. AORTA has a low-granularity model of turn conflict. If two
vehicles at any point along two different turns could ever come into contact,
then the turns are always in conflict, no matter where the agents actually are
along that turn. This is visualized in figure \ref{fig:conflicts}. The
alternative of tiling intersections \cite{JAIR08-dresner} would provide more
granularity, but it adds computational complexity and would be impractical for
many of the complex intersection geometries inferred from OSM.

@@

\begin{figure}[h]
  \centering \includegraphics[scale=0.5]{turn_conflicts.png}
  \caption{An agent cannot turn left while agents on the perpendicular road cross}
  \label{fig:conflicts}
  \vspace{-15pt}
\end{figure}

@@

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. Figure \ref{fig:gridlock} demonstrates this.
AORTA can detect this by finding cycles in the graph of the ``is following''
relation.

\begin{figure}[h]
  \centering \includegraphics[scale=0.25]{gridlock.png}
  \caption{Each of the red agents wants to turn left, so none of them
           can. The traffic jam will slowly build up around this intersection.}
  \label{fig:gridlock}
  \vspace{-10pt}
\end{figure}

@@

Each policy has parameters, so the configuration used is relevant. Stop signs
accept agents FCFS, refusing agents until they have idled at $0$ speed for $1$
second. This delay approximates human reaction time. The traffic signal policy
requires timings and groups of simultaneous turns for each intersection. The
configuration tested here uses a breadth-first search from one major road's
lanes to try to synchronize green signals for the agents near that lane to not
stop for several intersections. This heuristic needs improvement in how it
handles the remaining groups of turns that are scheduled after the first
sequence. Finally, the autonomous reservation policy allows new agents to join
existing groups of agents if their turn is compatible with the others. To
prevent one heavy direction of traffic from hogging the intersection
indefinitely, a timeout preemptively cycles through reservations. Ideally this
should be based on the relative numbers of agents depending on their
group of reservations.

@@

\begin{figure}[h]
  \centering \includegraphics[width=\linewidth]{avg_lag_agent_atx.png}
  \caption{The distribution of average delay reveals that traffic signals can
           cause the most lag.}
  \label{fig:avg_lag}
  \vspace{-10pt}
\end{figure}

If an intersection immediately permits an agent to proceed and grants other
agents in the same lane similar access (because this new agent must still follow
these others), then the new agent may simply proceed at full speed limit without
pausing. This gives the least possible delay spent along some lane, so any extra
time is defined as lag and is used to measure intersection performance. Figure
\ref{fig:avg_lag} graphs the average of this delay per agent. On average,
reservation policies cause less than $1$ second of delay at each intersection,
stop signs cause about $7$ seconds lag, and traffic signals cause $16$ seconds.

\begin{figure}[h]
  \centering \includegraphics[width=\linewidth]{agent_cnt_atx.png}
  \caption{Autonomous reservations outperform stop signs, and traffic signal
           gridlocked.}
  \label{fig:agent_cnt}
  \vspace{-10pt}
\end{figure}

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. Figure
\ref{agent_cnt} reveals that autonomous reservations let agents reach their
destination twice as quickly as the competition, with a steady-state of about
$400$ agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.
